Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-10-15 main.tex
intro & abstract
conclusion

We follow the approach of~\cite{Mohri03ijfcs} for the computation of distances
between words and languages
and extend it to infinite alphabets.
% with models of symbolic weighted automata and transducers.
%
The models presented in this section generalize weighted automata and transducers~\cite{Droste09handbook}
%over finite alphabets, see \eg~\cite{Mohri03ijfcs},
by labeling each transition with a weight function
that takes the input and output symbols as parameters.
These functions generalize the guards 
of symbolic automata~\cite{Veanes12symbolic,dAntoniVeanes17CAV,dAntoni21CACM},
from Boolean domains to commutative semirings.
%
They are particular cases of the very general model of 
Weighted Symbolic Automata with Data Storage~\cite{Herrmann16dlt}.

@@

\noindent
The state entering function of $B_{A, T}$ is defined
for all $\<p_1, q_1> \in Q'$, %$p_1 \in P$, $q_1 \in Q$
by:
%
\begin{equation} \label{eq:initprime}
\init'\bigl(\< p_1, q_1>\bigr) = \init_A(p_1) \otimes \init_T(q_1).
\end{equation}
%
The transition function $\wei'_1$ of $B_{A, T}$
will roughly perform a synchronized product of transitions defined by $\wei_1$
($A$ reading in input word $s$),
$\wei_{01}$ ($T$ reading in output word $t$ and not in input word $s$),
and $\wei_{11}$ ($T$ reading both in input word $s$ and in output word $t$).
%
Moreover, the $\varepsilon$-transitions to eliminate come from the simulation, 
by $\wei'_1$, of the transitions defined by $\wei_1$ and $\wei_{10}$ 
($A$ and $T$ reading in input word $s$ and not in output word $t$).
%
% Initialize state leaving functions i
%and $\final'(p, q) = \final_A(p) \otimes \final_T(q)$.
%
%Every transition of $B_{A, T}$ will
%simulate a sequence of transitions of $T$ performing the following steps:
%advance in the input word while staying immobile in the output word,
%and then make one step in the output word (and advance in the input word or not).
%
\noindent
Let us construct the transition function $\wei'_1$ iteratively.
%
\noindent
Initially, for all $p_1, p_2 \in P$, and $q_1, q_2 \in Q$, let
%
\begin{align}
%\begin{array}{rcl}
\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_2>\bigr) & = 
\bigl({\displaystyle\bigoplus_{p_1 = p_2}} \wei_{01}(q_1, q_2)\;\bigr)
\oplus
{\bigoplus}_\Sigma^1 \bigl(\wei_1(p_1, p_2) \otimes_1 \wei_{11}(q_1, q_2)\bigr)
\label{eq:init-wei}\\
%
\final'\bigl(\<p_1, q_1>\bigr) & =  \final_A(p_1) \otimes \final_T(q_1)
\label{eq:init-final}
%\end{array}
\end{align}
We recall that by convention, 
${\displaystyle\bigoplus_{p_1 = p_2}} \wei_{01}(q_1, q_2)$
is equal to $\zero$ if $p_1 \neq p_2$.

@@

We studied 3 properties of Symbolic Weighted language models, %over infinite alphabets,
and applied them to the problem of parsing
with infinitely many possible input symbols (typically timed events).
Our approach extends conventional parsing and weighted parsing
by computing a derivation tree modulo
a distance between words defined by a SW transducer given in input.
This allows to consider finer word relationships than strict equality.

This work can be extended in several directions. 
First,  the best search algorithm could be generalized from $1$-best to $n$-best~\cite{Huang05kbest},
and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}). Second,
%we believe that
the complexity bounds of the algorithms could be more precisely characterized,
as well as expressiveness of \swM compared to the automata of 
\eg~\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.
Finally, the  best search algorithm presented here works offline, whereas
an on-the-fly automata construction would allow for online parsing,
a suitable feature in the context of applications such as 
\eg automatic music transcription.

\florent{comparison of expressiveness with the automata 
of\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.}

@@

>>>>

We follow the approach of~\cite{Mohri03ijfcs} for the computation of distances
between words and languages
and extend it to infinite alphabets.
% with models of symbolic weighted automata and transducers.
%
The models presented in this section generalize weighted automata and transducers~\cite{Droste09handbook}
%over finite alphabets, see \eg~\cite{Mohri03ijfcs},
by labeling each transition with a weight function
that takes the input and output symbols as parameters.
These functions generalize the guards 
of symbolic automata~\cite{Veanes12symbolic,dAntoniVeanes17CAV,dAntoni21CACM},
from Boolean domains to commutative semirings.
%
They are particular cases of the very general model of 
Weighted Symbolic Automata with Data Storage~\cite{Herrmann16dlt}. 
\florent{review 3: = classical transducers}

@@

\noindent
The state entering function of $B_{A, T}$ is defined
for all $\<p_1, q_1> \in Q'$, %$p_1 \in P$, $q_1 \in Q$
by:
%
\begin{equation} \label{eq:initprime}
\init'\bigl(\< p_1, q_1>\bigr) = \init_A(p_1) \otimes \init_T(q_1).
\end{equation}
%
The transition function $\wei'_1$ of $B_{A, T}$
will roughly perform a synchronized product of transitions defined by $\wei_1$
($A$ reading in input word $s$),
$\wei_{01}$ ($T$ reading in output word $t$ and not in input word $s$),
and $\wei_{11}$ ($T$ reading both in input word $s$ and in output word $t$).
%
Moreover, the $\varepsilon$-transitions to eliminate come from the simulation, 
by $\wei'_1$, of the transitions defined by $\wei_1$ and $\wei_{10}$ 
($A$ and $T$ reading in input word $s$ and not in output word $t$).
%
% Initialize state leaving functions i
%and $\final'(p, q) = \final_A(p) \otimes \final_T(q)$.
%
%Every transition of $B_{A, T}$ will
%simulate a sequence of transitions of $T$ performing the following steps:
%advance in the input word while staying immobile in the output word,
%and then make one step in the output word (and advance in the input word or not).
%
\noindent
Let us construct the transition function~$\wei'_1$ iteratively.
%
\noindent
Initially, for all $p_1, p_2 \in P$, and $q_1, q_2 \in Q$, let
%
\begin{align}
%\begin{array}{rcl}
\wei'_1\bigl( \< p_1, q_1>, \< p_2, q_2>\bigr) & = 
\bigl({\displaystyle\bigoplus_{p_1 = p_2}} \wei_{01}(q_1, q_2)\;\bigr)
\oplus
{\bigoplus}_\Sigma^1 \bigl(\wei_1(p_1, p_2) \otimes_1 \wei_{11}(q_1, q_2)\bigr)
\label{eq:init-wei}\\
%
\final'\bigl(\<p_1, q_1>\bigr) & =  \final_A(p_1) \otimes \final_T(q_1)
\label{eq:init-final}
%\end{array}
\end{align}
We recall that by convention, 
${\displaystyle\bigoplus_{p_1 = p_2}} \wei_{01}(q_1, q_2)$
is equal to $\zero$ if $p_1 \neq p_2$.

@@

We studied properties of three Symbolic Weighted language models, %over infinite alphabets,
and applied them to the problem of parsing
with infinitely many possible input symbols (typically timed events).
One originality of this approach 
%extends conventional parsing and weighted parsing
is to compute derivation trees modulo
a distance between words defined by a given SW transducer.
This allows to consider finer word relationships than strict equality.

This work can be extended in several directions. 
The best search algorithm could be generalized from $1$-best to $n$-best~\cite{Huang05kbest},
and to $k$-\emph{closed} semirings~\cite{Mohri02semiring}
(instead of \emph{bounded}, which corresponds to $0$-\emph{closed}). 
The complexity bounds of the best-search algorithms 
could also certainly be improved.
%as well as expressiveness of \swM compared to the automata of 
%\eg~\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.
Another possible study could be the existence of best-search algorithms
for general models like~\cite{Herrmann16dlt}.
Finally, the  best search algorithm presented here works offline, 
whereas an on-the-fly approach coupling automata construction and best-search 
would be interesting \eg for online XML validation or filtering.
%a suitable feature in the context of applications such as 
%\eg automatic music transcription.

%\florent{comparison of expressiveness with the automata 
%of\cite{Segoufin06csl,KaminskiFrancez94,NevenSchwentickVianu04FSMinfinite,Bojanczyk11FO2}.}

@@

\section{Proof of Proposition~\ref{algo-shortest}}\label{app:shortpath} % Shortest path
\input{algo-best-lemma.tex}
